---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Habitat classification data from Lumbierres *et al.* 2021

[![lifecycle](https://img.shields.io/badge/Lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![DOI](https://img.shields.io/badge/DOI-TODO-blue.svg)](https://doi.org/TODO)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/README-", fig.align = "center")
```

This repository contains code to generate global habitat classification data (based on Lumbierres *et al.* 2021). Specifically, source data are downloaded from the [Zenodo Digital Archive](https://doi.org/10.5281/zenodo.5146072), and then projected to the World Behrmann coordinate reference system using (100 m resolution; [ESRI:4326 coordinate reference system](https://epsg.io/54017)). A processed version of the data can be downloaded from the [Zenodo Digital Repository](https://zenodo.org/badge/latestdoi/TODO). **All credit should go to Lumbierres *et al.* (2021), please cite their work if you use this dataset.**

```{r "preliminary_processing", echo = FALSE, message = FALSE, warning = FALSE}
# load packages
library(dplyr)
library(terra)
library(ggplot2)

# find file paths
data_path <- dir("results", "^.*tif$", full.names = TRUE)

# load raster
data_raster <- terra::rast(data_path)

# extract data to plot (note that we spatially sample data to reduce run time)
plot_data <-
  data_raster %>%
  terra::spatSample(size = 400000, method = "regular", as.raster = TRUE) %>%
  terra::as.data.frame(xy = TRUE) %>%
  dplyr::rename(value = 3) %>%
  dplyr::mutate(title = "Habitat classification") %>%
  dplyr::mutate(subtitle = basename(data_path)) %>%
  dplyr::mutate(value = as.character(value))

# create an underlying map
p <-
  ggplot() +
  coord_fixed(expand = FALSE) +
  geom_tile(
    data = plot_data,
    mapping = aes(x = x, y = y, fill = value)
  ) +
  facet_wrap(~ title + subtitle) +
  theme(
    legend.pos = "hide",
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "null"),
    axis.ticks.margin = unit(0, "null"),
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid = element_blank(),
    legend.text = element_text(size = 8),
    legend.margin = unit(0, "null"),
    plot.margin = unit(c(0.001, 0.001, 0.001, 0.01), "pt"),
    strip.background = ggplot2::element_rect(color = "black", fill = "black"),
    strip.text = ggplot2::element_text(color = "white")
  )
```

```{r "map", echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.4, fig.width = 5.0, dpi = 250}
# render plot
print(p)
```

## Metadata

The habitat classification data are stored in a raster (GeoTIFF) file. The file name contains the version number of the original datasets (i.e., per the digital object identifier). The latest version of the file is ``r basename(data_path)``.

The metadata for the ``r basename(data_path)`` raster file is shown below.

```{r, echo = FALSE, comment = NA}
withr::with_dir("results", {
  cat(
    system(paste("gdalinfo", basename(data_path)), intern = TRUE),
    sep = "\n"
  )
})
```

## Usage

You can run the code to produce the dataset using the system command `make clean all`. After processing has successfully completed, the code will produce two raster (GeoTIFF) files in the `results` directory.

## Software requirements

Operating system

* Ubuntu (21.10) (300 Gb disk space available)

Software

* GNU make (version 4.1)
* GDAL (version 3.2.2)
* R (version 4.1.1)

System packages

* `gdal-bin`
* `libudunits2-dev`
* `libgdal-dev`
* `libgeos-dev`
* `libproj-dev`
* `libxml2-dev`
* `libssl-dev`

## Citation

This dataset was derived from Lumbierres *et al.* (2021). If you use this dataset, you should cite their work - they deserve all the credit!

> Lumbierres, M (2021). *Map of habitat classes (Level 1) from the IUCN Habitat.* Zenodo Digital Repository. Available at https://doi.org/10.5281/zenodo.5146072.

> Lumbierres M, Dahal PR, Di Marco M, Butchart SHM, Donald PF, and Rondinini C (2021) Translating habitat class to land cover to map area of habitat of terrestrial vertebrates. *Conservation Biology*, **In press**, DOI:10.1111/cobi.13851. Available at https://doi.org/10.1111/cobi.13851.
